---
title: Charles Zheng EE 378b HW 2
output:
  html_document:
    mathjax: default
---

Charles Zheng EE 378b HW 3
========================================================

# Setup

Loading the data:

```{r}
library(parallel)
library(rARPACK)
library(magrittr)
library(graph)
list.files("ee378b")
temp <- as.matrix(read.table("ee378b/CA-GrQc.txt", sep = "\t", skip = 4))
apply(temp, 2, min)
apply(temp, 2, max)
A <- ftM2adjM(temp)
diag(A) <- 0
temp <- apply(A, 1, sum) > 0
A <- A[temp, temp]
dim(A) # 5241 5241
D <- diag(apply(A, 1, sum))
```

Define the following functions:
```{r}
modularity_mat <- function(A) {
  d <- apply(A, 1, sum)
  m2 <- sum(d)
  A - 1/m2 * d %*% t(d)
}

modularity <- function(B, clusters) {
  k <- length(unique(clusters))
  mod <- 0
  for (i in unique(clusters)) {
    mod<- mod + sum(B[clusters == i, clusters == i])
  }
  mod
}

mod_search <- function(B, vec, inds, all_values = FALSE, mc = 0) {
  svec <- sort(vec)
  ff <- function(ind) modularity(B, vec > svec[ind])
  if (mc > 0) {
    mods <- unlist(mclapply(inds, ff, mc.cores = mc))
  } else {
    mods <- sapply(inds, ff)  
  }
  if (all_values) return(mods)
  vec > (svec[mods == min(mods)])
}

eigencoords <- function(m, k = 2) {
  res <- eigs(m, k)
  if (k == 1) return(res$vectors[, 1])
  apply(res$vectors[, 1:k], 1, function(v) v/sqrt(sum(v^2)))
}
```

Compare how well laplacian, normalized laplacian, and modularity matrix perform for two-cluster case.

```{r fig.width=7, fig.height=6}
B <- modularity_mat(A)
n <- dim(A)[1]
search_inds <- floor(seq(n/200, n, n/100))
isqrtD <- diag(1/sqrt(diag(D)))
L <- A - D
N <- isqrtD %*% A %*% isqrtD - diag(rep(1, n))
L_coords <- eigencoords(L, 1)
plot(mod_search(B, L_coords, search_inds, TRUE, 6))
title("Laplacian")
```

Normalized laplacian
```{r fig.width=7, fig.height=6}
N_coords <- eigencoords(N, 1)
plot(mod_search(B, N_coords, search_inds, TRUE, 6))
title("Normalized laplacian")
```

Modularity matrix
```{r fig.width=7, fig.height=6}
B_coords <- eigencoords(B, 1)
plot(mod_search(B, B_coords, search_inds, TRUE))
title("Modularity matrix")
```

# Iterative Proceure

```{r}
mat <- B
min_dim <- 10
current_inds <- 1:20
search_K <- 5

catn <- function(...) { cat(...); cat('\n') }

itersplit <- function(B, mat, search_K, min_dim = 20, current_inds = 1:dim(B)[1],
                      diagnostic = FALSE, pref = "") {
  if (length(current_inds) < 2 * min_dim) {
    if (diagnostic) catn(paste0(pref, ": final size :", length(current_inds)))    
    return(list(current_inds))
  }
  if (diagnostic) catn(paste0(pref, ": NEW :", length(current_inds)))
  smat <- mat[current_inds, current_inds]
  sB <- B[current_inds, current_inds]
  coords <- eigencoords(smat, 1)
  sn <- length(current_inds)
  search_inds <- floor(seq(min_dim, sn - min_dim, sn/search_K))
  mods <- mod_search(sB, coords, search_inds, TRUE)
  #if (diagnostic) catn(paste0(pref, paste(mods, collapse = " ")))
  mod0 <- sum(sB)
  if (max(mods) < mod0) {
    if (diagnostic) catn(paste0(pref, ": Sum too small"))
    return (list(current_inds))
  } else {
    if (diagnostic) catn(paste0(pref, ": Mod improvement ", mod0, " to ", max(mods)))
    svec <- sort(coords)
    split <- coords > svec[search_inds[mods == max(mods)]][1]
    ans1 <- itersplit(B, mat, search_K, min_dim, current_inds[split], diagnostic, paste0(pref, "-"))      
    ans2 <- itersplit(B, mat, search_K, min_dim, current_inds[!split], diagnostic, paste0(pref, "*"))
    ans <- c(ans1, ans2)
    if (pref != "") return (ans)
  }
  # only in root
  clusters <- 1:dim(B)[1]
  for (i in 1:length(ans)) clusters[ans[[i]]] <- i
  clusters
}

Bclusters <- itersplit(B, B, 20, diagnostic = FALSE)
modularity(B, Bclusters)
Lclusters <- itersplit(B, L, 20, diagnostic = FALSE)
modularity(B, Lclusters)
Nclusters <- itersplit(B, N, 20, diagnostic = FALSE)
modularity(B, Nclusters)
```

# Greedy stuff.

Quickly (?) compute change in modularity

```{r}
# computes change in modularity from switching ind to label
mod_change <- function(B, clusters, ind, lbl) {
  others <- rep(TRUE, dim(B)[1]) %>% {.[ind] <- FALSE; .}
  new_neighbors <- others & (clusters == lbl)
  old_neighbors <- others & (clusters == clusters[ind])
  2 * (sum(B[ind, new_neighbors]) - sum(B[ind, old_neighbors]))
}

# test the method
clusters <- Bclusters
ind <- 2
lbl <- 2
modularity(B, clusters)
modularity(B, clusters) + mod_change(B, clusters, ind, lbl)
modularity(B, clusters %>% {.[ind] <- lbl; .})
```

Try one (big) iteration of greedy algorithm.

```{r}
greedo <- function(B, clusters, n_its = 1) {
  n <- dim(B)[1]
  k <- max(clusters)
  for (it in 1:n_its) {
    for (i in 1:n) {
      bonds <- sapply(1:k, function(j) sum(B[i, clusters == j]))
      clusters[i] <- order(-bonds)[1]
    }
  }
  clusters
}

Bclusters2 <- greedo(B, Bclusters, 2)
modularity(B, Bclusters)
modularity(B, Bclusters2)
```

Try with random intializations.
```{r}
K_cluster <- 100
rand_inits <- lapply(1:10, function(v) {set.seed(v); sample(K_cluster, n, TRUE) })
updated <- mclapply(rand_inits, function(v) greedo(B, v, 10), mc.cores = 5)
```

