---
title: "Random images misclassification rate"
output: html_document
---

Model: consider a spherical parameter space with radius $R$, $\Omega = B_R(0) \in \mathbb{R}^d$.

Let $P$ be a distribution with density $p(x)$ supported on $\Omega$.

Generate $k$ means, $\mu_i \sim P$. (Each mean belongs to an image.)  Let $\sigma$ be a noise parameter.

Generate $n$ observations per mean, $x_{i, 1},\hdots, x_{i, n} \sim N(\mu_i, \sigma^2 I)$.

From $n_{tr}$ observations per mean estimate $\hat{\mu}_i$ for $i = 1,\hdots, k$.

Then assess the misclassification rate on the remaining $k(n-n_{tr})$ test points.

Set parameters:

```{r}
# PARAMETERS
# radius of the parameter ball
omega_radius <- 2
# dimension of the parameter space
omega_dim <- 2
# noise
sigma <- 0.1
# number of images
k <- 10
# number of total observations
n <- 10
# number of training observations
n_tr <- 5
# density function p(x)
p <- function(x) {
  
}
```

Generate random means:

```{r}
mus <- 
```

Mean rates for each angle
```{r, fig.width=6, fig.height=3}
plot(1:5)
```

