Charles Zheng Homework 3
========================================================

## Computation

Consider a CAR model, with a weight matrix $W$ which is not necessarily symmetric and has zero diagonal.
Let $\tilde{W}$ denote the lower triangle of $W$.
We have
$$
\frac{p(\epsilon)}{p(0)} = \exp\left(-\frac{1}{2\tau^2} \epsilon^T (I - 2\phi \tilde{W}) \epsilon\right)
$$
which is equivalent to
$$
\frac{p(\epsilon)}{p(0)} = \exp\left(-\frac{1}{2\tau^2} \epsilon^T (I - \phi (\tilde{W} + \tilde{W}^T)) \epsilon\right).
$$
Since $\tilde{W} + \tilde{W}^T$ is symmetric, we might as well assume that $W$ is symmetric.
Hence
$$
\epsilon \sim N(0, \tau^2(I - \phi W)^{-1})
$$
Now suppose
$$
W = V D V^T.
$$
Then
$$
\tau^2 (I - \phi W)^{-1} = \tau^2 V (I - \phi D)^{-1} V^T 
$$

## Model

Since we have Poisson data, we use the model
$$
y \sim Poisson(e^{X\beta + \epsilon})
$$
where
$$
\epsilon \sim N(0, \tau^2(I - \phi W)^{-1})
$$
Here the model is no longer a CAR model due to the Poisson transformation.
Instead, $\epsilon$ act as latent variables which induce spatial correlation.

```{r, warning=FALSE, message=FALSE}
library(maptools)
library(RColorBrewer)
library(classInt)
library(magrittr)
library(spdep)
library(pracma)
```

## Loading the data

```{r}
shapes <- readShapeSpatial("scotlip/scotlip.shp")
plot(shapes)
nb <- poly2nb(shapes)
W <- nb2mat(nb, zero.policy=TRUE)
W2 <- (W * lower.tri(W)) %>% {. + t(.)}
V <- res$vectors
d <- res$values
y <- shapes@data$CANCER
X <- shapes@data %$% cbind(POP, CEXP, AFF, AREA, PERIMETER)
X[, "POP"] <- log(X[, "POP"])
X[, "CEXP"] <- log(X[, "CEXP"])
X[, "AREA"] <- log(X[, "AREA"])
X[, "PERIMETER"] <- log(X[, "PERIMETER"])
dim(X)
n <- dim(W)[1]
```

## Models

Model file

```
data {
  for (i in 1:n) {
    zeros[i] <- 0
  }
}

model {
  phi ~ dunif(0, 0.9)
  tau ~ dgamma(2, 100)
  tau2 ~ dgamma(2, 100)
  beta0 ~ dnorm(0, tau2)
  for (i in 1:p) {
    beta[i] ~ dnorm(0, tau2)
  }
  eps[] ~ dmnorm(zeros[], T[ , ])
  for (i in 1:n) {
    mu[i] <- exp(inprod(X[i, 1:p], beta[]) + beta0 + eps[i])
    y[i] ~ dpois(mu[i])
  }
  for (i in 1:n) {
    for (j in 1:n) {
      Wt[i,j] <- pow(tau, 2)/(1 - phi * d[j]) * Vt[i, j]
    }
  }
  for (i in 1:n) {
    for (j in 1:n) {
      T[i, j] <- inprod(V[i,1:n],Wt[1:n,j])
    }
  }
}
```

Intializing model...

```{r, results='hide'}
model <- jags.model("pois_sp.bug", data=list(eps = rnorm(56), d = d, n = 56,
                                             V = V, Vt = t(V), y = y, X = X, p = 5),
                    inits = list(phi = 0.2, tau = 2, tau2 = 2,
                                 beta0 = 0, beta = rep(0, 5)), n.chains = 5)
```

Draw samples

```{r, results='hide'}
out <- coda.samples(model, c("phi", "beta", "eps"), n.iter = 2000)
```

```{r}
traceplot(out[[1]][, "beta[1]"])
```





