---
title: "Robustness of Knockoffs in Realistic Settings"
author: "Charles Zheng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Introduction

The knockoff filter (Barber and Candes 2014) is a method for controlling false discovery rate for
linear regression coefficients.

## False discovery rate

The Type I error criteria used in Barber and Candes (2014) is the false discovery rate.
The criteria is defined as

$$
FDR = E[V/\max(1,R)]
$$

where $V$ is the number of false rejections (negatives rejected) and $R$ is the total number of rejections.
Therefore we should see in practice that applying the knockoff filter with FDR threshold $q$,
over many datasets $(X^{(1)}, Y^{(1)}), ... , (X^{(m)}, Y^{(m)})$, the counts of false rejections $V^{(i)}$ and total rejections $R^{(i)}$ should satisfy
$$
q > \frac{1}{m}\sum_{i=1}^m \frac{V^{(i)}}{\max(1, R^{(i)})}
$$
or 
$$
q \approx \frac{1}{m}\sum_{i=1}^m \frac{V^{(i)}}{\max(1, R^{(i)})}
$$

## Validation

The knockoff filter was derived under a number of assumptions which are unlikely to hold in real data, namely:

- Linearity of the expectation of the respose in relation to the covariates
- Gaussianity of the noise
- Homoscedasticity and independence of the noise

Therefore, it is important to know how the Type I error of the procedure is affected in realistic settings
where these assumptions are violated.

A major obstacle to validating any inference procedure in real data is that the "true" regression coefficients are unknown.  Therefore, it is difficult to tell when actual Type I errors are made.  In light of this difficulty, we resort to creating artificial variables and including them in the inference procedure as known negatives

## Prostate Data

We separate the covariates and response

```{r}
library(reginference)
data(prostate)
x <- prostate[, 1:8]
y <- prostate[, 9]
```

We create artificial covariates.
NOTE: The knockoff procedure itself also creates artificial knockoff 'covariates'.
The variables created in this step are NOT to be confused with the knockoffs created by the knockoff procedure itself.
These variables are known negatives which are used to externally validate the knockoff procedure.

```{r}
x_false <- noised_projections(x, sigma = 0.1, q = 20)
x_aug <- cbind(x, x_false)
```

Now apply the knockoff procedure with FDR $q = 0.2$:
```{r}
sel <- knockoff::knockoff.filter(x_aug, y, fdr = 0.2)$selected
```

Compute the fdp
```{r}
sum(names(sel) %in% colnames(x_false))/max(1, length(sel))
```

### Experiments

The following function performs the experiment with given FDR-thresholds, number of artificial variates, and noise level of aritificial variates, and returns the average FDP over repeated trials.

```{r}
knockoff_experiment <- function(x, y, fdr, sigma, n_artificial, ntrials) {
  nq <- length(fdr)
  results <- matrix(0, ntrials, nq)
  for (i in 1:ntrials) {
    x_false <- noised_projections(x, sigma = sigma, n_artificial)
    x_aug <- cbind(x, x_false)
    for (j in 1:nq) {
      sel <- knockoff::knockoff.filter(x_aug, y, fdr = fdr[j])$selected
      results[i, j] <- sum(names(sel) %in% colnames(x_false))/max(1, length(sel))      
    }
  }
  results
}
```

Plot the nominal FDR versus the ground truth FDR (based on artificial variates).

```{r, fig.show='hold'}
fdr <- (1:10)/20
res <- knockoff_experiment(x, y, fdr, 0.1, 30, 20)
ave_fdr <- apply(res, 2, mean)
plot(fdr, ave_fdr)
abline(0, 1)
```

## Galaxy Data

```{r}
data(galaxy)
x <- galaxy[, 1:4]
y <- galaxy[, 5]
```

The knockoff filter fails to control FDR for the galaxy data.

```{r, fig.show='hold'}
fdr <- (1:10)/20
res <- knockoff_experiment(x, y, fdr, 0.1, 30, 20)
ave_fdr <- apply(res, 2, mean)
plot(fdr, ave_fdr, ylim = c(0, 1))
abline(0, 1)
```

Could the culprit be nonlinearity? Plot the fitted values of y versus y.

```{r, fig.show='hold'}
library(magrittr)
res <- lm(y ~ as.matrix(x))
plot(y, res$fitted)
summary(res)
```

Check the individual scatterplots

```{r, fig.show='hold'}
layout(matrix(1:4, 2, 2))
galaxy %$% plot(radial.position, velocity)
galaxy %$% plot(north.south, velocity)
galaxy %$% plot(east.west, velocity)
galaxy %$% plot(angle, velocity)
layout(1)
```

There is an interaction between radial position and east-west.
```{r, eval = FALSE}
library(rgl)
galaxy %$% plot3d(radial.position, east.west, velocity)
```

## Ozone data

```{r}
data(ozone)
names(ozone)
y <- ozone[, 1]
x <- ozone[, 2:9]
```

Plot the nominal FDR versus the ground truth FDR (based on artificial variates).

```{r, fig.show='hold'}
fdr <- (1:10)/20
res <- knockoff_experiment(x, y, fdr, 0.1, 30, 20)
ave_fdr <- apply(res, 2, mean)
plot(fdr, ave_fdr, ylim = c(0, 1))
abline(0, 1)
```

## More data

```{r}
library(alr4)
library(alr3)
data(UN2)
View(UN2)
y <- UN2[, 2]
x <- UN2[, -2]
```

```{r, fig.show='hold'}
fdr <- (1:10)/20
res <- knockoff_experiment(x, y, fdr, 0.1, 10, 200)
ave_fdr <- apply(res, 2, mean)
plot(fdr, ave_fdr, ylim = c(0, 1))
abline(0, 1)
title("UN data FDR")
#plot(y, x[, 1])
#plot(y, x[, 2])
#res <- lm(y ~ as.matrix(x))
#plot(y, res$fitted)
#library(rgl)
plot3d(x[, 1], x[, 2], y)
#plot(x[, 1], x[, 2])
#identify(x[, 1], x[, 2])
```

```{r, fig.show='hold'}
y <- UN2[-101, 2]
x <- UN2[-101, -2]
res <- knockoff_experiment(x, y, fdr, 0.1, 10, 200)
ave_fdr <- apply(res, 2, mean)
plot(fdr, ave_fdr, ylim = c(0, 1))
abline(0, 1)
title("UN without outlier")
```

```{r}
res <- loess(logFertility ~ logPPgdp + Purban, data = UN2)
help(loess)
range(UN2$logPPgdp)
range(UN2$Purban)
zg <- grid_design(UN2, c(1, 3))
yg <- predict(res, zg)
plot3d(zg[, 1], zg[, 3], yg)
```
